RILS-ROLS: Robust Symbolic Regression via Iterated Local Search and Ordinary Least Squares

### 2022-10-26 doc/paper.tex
Experimental section: working

SRBench is an open-source benchmarking project which merge a large set of diverse benchmark datasets, contemporary SR methods as well as ML methods around a shared model evaluation and the environment for analysis, see~\cite{la2021contemporary}. It brought researchers an easy handled environment to evaluation, compare, and analyze their methods. Among all benchmark sets, we involve two real-world inspired benchmark sets into our computational evaluation: 
\begin{itemize}
	\item \textsc{Feynman} benchmark set, inspired from physics and formulas/models that correspond to various natural laws.  
	It consists of $10^5$  samples, each described in~\cite{udrescu2020ai}. 
	\item \textsc{Strogatz} benchmark set from~\cite{la2016inference}. 
	Each dataset is represented by a 2-state system of first-order, ordinary differential equations. 
	The aim of each problem is to predict rate of change of the subsequent state w.r.t\  the current two states on which it depends. These equations describes various non-linear dynamics and exhibiting chaos.

@@

\subsection{Parameter tuning}
The \textsc{Rils}-\textsc{Rols} algorithms employs just two parameters that need to be tuned, $penalty_{size}$ and $tolerance_{error}$. The first parameter is set to 0.001(?), according to the preliminary experimental evaluation. The second parameter is tuned monitoring the two errors the algorithm produced on the training set; we choose $tolerance_{error}=?$ for which a largest average $R^2$ and the smallest average RMSE error over all instances from the training set are produced. Note that the training set is chosen randomly with a cardinalty of xx instances.

@@

\subsection{Comparison with other methods on real--world benchark sets}
In this section, we evaluate the performance of \textsc{Rils-Rols} algorithm with 15 other competitors from the literature that solve the problem of Symbolic regression. The results are reported in terms of three numerical Tables~\ref{tab:comp_noise0}--\ref{tab:comp_noise001}, where two real-world benchmark sets, \textsc{Feynman} and \textsc{Storgatz} are considered, utilizing  different values of the noise levels: 0.0 (no noise), 0.001 (a low-level noise), and 0.01 (a high-level noise), respectively. Concerning the added noise, the White Gaussian noise was added to the target ($y$-axis values) as a fraction of the signal RMS value, also applied in~\cite{la2021contemporary}, i.e. for target noise level $\alpha$, we have
$$ y_{noise} = y + \epsilon, \epsilon \sim \mathcal{N}\left(0, \alpha \sqrt{\frac{1}{N} \sum_{y_i^2}}\right). $$

>>>>

SRBench is an open-source benchmarking project which merge a large set of diverse benchmark datasets, contemporary SR methods as well as ML methods around a shared model evaluation and the environment for analysis, see~\cite{la2021contemporary}. It brought researchers an easy handled environment to evaluation, compare, and analyze their methods. Among all benchmark sets, we involve two real-world inspired benchmark sets into our computational evaluation: 
\begin{itemize}
	\item \textsc{Feynman} benchmark set, inspired from physics and formulas/models that correspond to various natural laws.  
	It consists of $10^5$  samples, each described in~\cite{udrescu2020ai}. Some exact models (equations) of problem instances from this benchmark are shown in Table~\ref{tab:Feynamn-Eq}.

   \begin{table}
		\centering
		\begin{tabular}{l}   \hline
	       $x = \sqrt{x_1^2 + x_2^2 - 2 x_1 x_2 \cos(\theta_1 - \theta_2)}$ \\
	       $ \theta_1 = \arcsin(n \sin \theta_2)$ \\
	       $E =  \frac{m c^2 }{1 - \frac{v^2}{c^2}}$ \\
	       $\omega = \frac{1 + \frac{v}{c}}{ \sqrt{1 - \frac{v^2}{c^2}}} \omega_0$ \\ \hline

	    \end{tabular}
		\caption{Some Feynman equations.}
		\label{tab:Feynamn-Eq}
	\end{table}

	\item \textsc{Strogatz} benchmark set from~\cite{la2016inference}. 
	Each dataset is represented by a 2-state system of first-order, ordinary differential equations. 
	The aim of each problem is to predict rate of change of the subsequent state w.r.t\  the current two states on which it depends. These equations describes various non-linear dynamics and exhibiting chaos.  The equations for some of the datasets of this benchmark are shown in Table~\ref{table:strogatz-ODEs}.

	%table, an instance -- example

	\begin{table}
		\centering
		\begin{tabular}{ll} \\ \hline
			Bacterial representation &   $x' = 20 - x - \frac{x \cdot y}{1 + 0.5 x^2 }$ \\ 
			                         &   $y' = 10 - \frac{x \cdot y}{1 + 0.5 x^2  }$ \\ \hline
			Shear Flow               &  $\theta' = \cot(\phi)\cos(\theta)$ \\
			                         &  $ \phi'  = ( \cos^2(\phi) + 0.1 \cdot \sin^2 (\phi)) \sin(\theta) $ \\ \hline
		\end{tabular}
	    \caption{Some Strogatz ODE problems.}
	    \label{table:strogatz-ODEs}
	\end{table}

@@

\subsection{Parameter tuning}
The \textsc{Rils}-\textsc{Rols} algorithms, apart of the most other methods for SR, employs just two parameters that need to be tuned, $penalty_{size}$ and $tolerance_{error}$. The first parameter is set to 0.001(?), according to the preliminary experimental evaluation. The second parameter is tuned monitoring the two errors the algorithm produced on the training set; we choose $tolerance_{error}=?$ for which a largest average $R^2$ and the smallest average RMSE error over all instances from the training set are produced. Note that the training set is chosen randomly with a cardinalty of xx instances.

@@

\subsection{Comparison with other methods on real--world benchark sets}
In this section, we evaluate the performance of \textsc{Rils-Rols} algorithm with 15 other competitors from the literature that solve the problem of Symbolic regression. The results are reported in terms of three numerical Tables~\ref{tab:comp_noise0}--\ref{tab:comp_noise001}, where two real-world benchmark sets, \textsc{Feynman} and \textsc{Storgatz} are considered, utilizing  different values of the noise levels: 0.0 (no noise), 0.001 (a low-level noise), and 0.01 (a high-level noise), respectively. Concerning the added noise, the White Gaussian noise was added to the target ($y$-axis values) as a fraction of the signal RMS value, also applied in~\cite{la2021contemporary}, i.e. for target noise level $\alpha$, we have
$$ y_{noise} = y + \epsilon, \epsilon \sim \mathcal{N}\left(0, \alpha \sqrt{\frac{1}{N} \sum {y_i^2}}\right). $$
